### 2019-09-24 source/cadabra/metric.tex
Initial commit
This is where it all begins ...

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/metric-inv.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{The inverse metric tensor in Riemann normal coordinates}

Here we calculate the Riemann normal expansion of the inverse metric, $g^{ab}$, by developing
the recursive sequences
\begin{align}
\label{eq:pdgab}
g^{ab}{}_{,d\ue} &= - \left(g^{cb}\Gamma^{a}{}_{c(d}\right){}_{,\ue)}
                    - \left(g^{ac}\Gamma^{b}{}_{c(d}\right){}_{,\ue)}\\[10pt]
\label{eq:pdGamma}
(n+3)\Gamma^a{}_{d(b,c\ue)} &= (n+1)\left(R^a{}_{(bc\Dot d,\ue)} - \left(\Gamma^a{}_{f(c}\Gamma^f{}_{b{\Dot d}}\right){}_{,\ue)}\right)
\end{align}
for $n=1,2,3,\cdots$. Note in these equations that the (extended) index $\ue$ contains $n$
normal indices.

We then construct a Taylor series for the metric using
\begin{dmath*}[spread=5pt]
g^{ab}(x) = g^{ab} + g^{ab}{}_{,c}x^c + \frac{1}{2!} g^{ab}{}_{,cd}x^cx^d + \frac{1}{3!} g^{ab}{}_{,cde}x^cx^dx^e + \cdots
          = g^{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g^{ab}{}_{,\uc}\>x^{.\uc}
\end{dmath*}

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $g^{ab}$}

@@

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.200} \end{dmath*}
\end{dgroup*}

@@

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.303} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.305} \end{dmath*}
\end{dgroup*}

@@

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.404} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.404} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.403} \end{dmath*}
\end{dgroup*}

@@

\begin{align*}
   g_{ab}(x) &=   g^{ab}
                + g^{ab}{}_{,c} x^c
                + \frac{1}{2!} g^{ab}{}_{,cd} x^c x^d
                + \frac{1}{3!} g^{ab}{}_{,cde} x^c x^d x^e +  \BigO{\eps^4}\\
             &= \cdb{metric4.501} + \BigO{\eps^4}
\end{align*}

@@

   def flatten_Rabcd (obj):
       substitute (obj,$R^{a}_{b c d}   -> g^{a e} R_{e b c d}$)
       substitute (obj,$R_{a}^{b}_{c d} -> g^{b e} R_{a e c d}$)
       substitute (obj,$R_{a b}^{c}_{b} -> g^{c e} R_{a b e d}$)
       substitute (obj,$R_{a b c}^{d}   -> g^{d e} R_{a b c e}$)
       unwrap     (obj)
       sort_product   (obj)
       rename_dummies (obj)
       return obj

@@

   def add_tags (obj,tag):
      n = 0
      ans = Ex('0')
      for i in obj.top().terms():
         foo = obj[i]
         bah = Ex(tag+'_{'+str(n)+'}')
         ans := @(ans) + @(bah) @(foo).
         n = n + 1
      return ans

   def clear_tags (obj,tag):
      ans := @(obj).
      foo  = Ex(tag+'_{a?} -> 1')
      substitute (ans,foo)
      return ans

@@

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $g^{ab}$}

@@

   # symmetrised partial derivatives of g^{ab}

   gab00:=g^{a b}.                                              # cdb (gab00.101,gab00)

   gab01:= - g^{c b}\Gamma^{a}_{c d} - g^{a c}\Gamma^{b}_{c d}. # cdb (gab01.101,gab01)

   gab02:=\partial_{e}{ @(gab01) }.                             # cdb (gab02.101,gab02)
   distribute   (gab02)                                         # cdb (gab02.102,gab02)
   product_rule (gab02)                                         # cdb (gab02.103,gab02)
   substitute   (gab02, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab02.104,gab02)
   distribute   (gab02)                                         # cdb (gab02.105,gab02)

   gab03:=\partial_{f}{ @(gab02) }.                             # cdb (gab03.101,gab03)
   distribute   (gab03)                                         # cdb (gab03.102,gab03)
   product_rule (gab03)                                         # cdb (gab03.103,gab03)
   substitute   (gab03, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab03.104,gab03)
   distribute   (gab03)                                         # cdb (gab03.105,gab03)

   gab04:=\partial_{g}{ @(gab03) }.                             # cdb (gab04.101,gab04)
   distribute   (gab04)                                         # cdb (gab04.102,gab04)
   product_rule (gab04)                                         # cdb (gab04.103,gab04)
   substitute   (gab04, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab04.104,gab04)
   distribute   (gab04)                                         # cdb (gab04.105,gab04)

   gab05:=\partial_{h}{ @(gab04) }.                             # cdb (gab05.101,gab05)
   distribute   (gab05)                                         # cdb (gab05.102,gab05)
   product_rule (gab05)                                         # cdb (gab05.103,gab05)
   substitute   (gab05, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab05.104,gab05)
   distribute   (gab05)                                         # cdb (gab05.105,gab05)

@@

\begin{cadabra}
   # prepare first six terms in the Taylor series expansion of g^{ab}(x)

@@

   # ------------------------------------------------------------------------------------
   # this block only produces formatted output, it is not part of the main computation
   # ------------------------------------------------------------------------------------

@@

   metric:=@(term0)
         + (1/1) @(term1)
         + (1/2) @(term2)
         + (1/6) @(term3)
         + (1/24) @(term4)
         + (1/120) @(term5).  # cdb(metric.301,metric)

@@

   Xterm0 := @(term0).
   Xterm1 := (1/1) @(term1).     # zero
   Xterm2 := (1/2) @(term2).
   Xterm3 := (1/6) @(term3).
   Xterm4 := (1/24) @(term4).
   Xterm5 := (1/120) @(term5).

@@

   # Manipulating these expressions is hampered by the presence of the partial derivative on Rabcd.
   # Thus we can't freely rasie/lower indices on the dRabcd terms. But we can do so on the first
   # derivatives (since these are evaluated at x=0 where the connection vanishes).

   substitute       (Xterm2,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm2.301,Xterm2)
   substitute       (Xterm3,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm3.301,Xterm3)
   substitute       (Xterm4,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm4.301,Xterm4)
   substitute       (Xterm5,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm5.301,Xterm5)

   substitute       (Xterm2,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm2.302,Xterm2)
   substitute       (Xterm3,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm3.302,Xterm3)
   substitute       (Xterm4,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm4.302,Xterm4)
   substitute       (Xterm5,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm5.302,Xterm5)

   substitute       (Xterm2,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm2.303,Xterm2)
   substitute       (Xterm3,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm3.303,Xterm3)
   substitute       (Xterm4,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm4.303,Xterm4)
   substitute       (Xterm5,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm5.303,Xterm5)

   substitute       (Xterm2,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm2.304,Xterm2)
   substitute       (Xterm3,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm3.304,Xterm3)
   substitute       (Xterm4,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm4.304,Xterm4)
   substitute       (Xterm5,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm5.304,Xterm5)

   sort_product     (Xterm2)  # cdb(Xterm2.305,Xterm2)
   sort_product     (Xterm3)  # cdb(Xterm3.305,Xterm3)
   sort_product     (Xterm4)  # cdb(Xterm4.305,Xterm4)
   sort_product     (Xterm5)  # cdb(Xterm5.305,Xterm5)

   rename_dummies   (Xterm2)  # cdb(Xterm2.306,Xterm2)
   rename_dummies   (Xterm3)  # cdb(Xterm3.306,Xterm3)
   rename_dummies   (Xterm4)  # cdb(Xterm4.306,Xterm4)
   rename_dummies   (Xterm5)  # cdb(Xterm5.306,Xterm5)

   canonicalise     (Xterm2)  # cdb(Xterm2.307,Xterm2)
   canonicalise     (Xterm3)  # cdb(Xterm3.307,Xterm3)
   canonicalise     (Xterm4)  # cdb(Xterm4.307,Xterm4)
   canonicalise     (Xterm5)  # cdb(Xterm5.307,Xterm5)

   # We can simplify Xterm2 and Xterm3 by careful juggling of the indices (swapping free indices on selected terms)

   tmp = add_tags (Xterm2,'\\mu')       # cdb (tmp.001,tmp)
   zoom (tmp, $\mu_{1} Q??$)            # cdb (tmp.002,tmp)
   substitute (tmp, $R^{b}_{c d}^{a} x^{c} x^{d} -> R^{a}_{c d}^{b} x^{c} x^{d}$)  # cdb (tmp.003,tmp)
   unzoom (tmp)
   Xterm2 = clear_tags (tmp,'\\mu')     # cdb (Xterm2.401,Xterm2)

   tmp = add_tags (Xterm3,'\\mu')       # cdb (tmp.011,tmp)
   zoom (tmp, $\mu_{1} Q??$)            # cdb (tmp.012,tmp)
   substitute (tmp, $\partial_{c}{R^{b}_{d e}^{a}} x^{c} x^{d} x^{e} -> \partial_{c}{R^{a}_{d e}^{b}} x^{c} x^{d} x^{e}$)  # cdb (tmp.013,tmp)
   unzoom (tmp)
   Xterm3 = clear_tags (tmp,'\\mu')     # cdb (Xterm3.401,Xterm3)

   Xterm0 = reformat_xterm (Xterm0,  1)    # cdb(Xterm0.308,Xterm0)
   Xterm2 = reformat_xterm (Xterm2,  3)    # cdb(Xterm2.308,Xterm2)
   Xterm3 = reformat_xterm (Xterm3,  6)    # cdb(Xterm3.308,Xterm3)
   Xterm4 = reformat_xterm (Xterm4,360)    # cdb(Xterm4.308,Xterm4)
   Xterm5 = reformat_xterm (Xterm5,360)    # cdb(Xterm5.308,Xterm5)

   # metric to 4th and 6th order terms in powers of x^a

@@

   end_stage_2 = time.time()
\end{cadabra}

@@

\begin{dgroup*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric3.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric4.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric5.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric6.301} \end{dmath*}
\end{dgroup*}

@@

\begin{cadabra}
   beg_stage_4 = time.time()

   # final housekeeping

   # lower the ^{ab} indices to _{uv}

   tmp0 := g_{a u} g_{b v} @(term0).
   tmp1 := g_{a u} g_{b v} @(term1).
   tmp2 := g_{a u} g_{b v} @(term2).
   tmp3 := g_{a u} g_{b v} @(term3).
   tmp4 := g_{a u} g_{b v} @(term4).
   tmp5 := g_{a u} g_{b v} @(term5).

   distribute           (tmp1)  # cdb(tmp1.501,tmp1)
   eliminate_metric     (tmp1)  # cdb(tmp1.502,tmp1)
   eliminate_kronecker  (tmp1)  # cdb(tmp1.503,tmp1)
   tmp1 = flatten_Rabcd (tmp1)
   canonicalise         (tmp1)  # cdb(tmp1.506,tmp1)

   distribute           (tmp2)  # cdb(tmp2.501,tmp2)
   eliminate_metric     (tmp2)  # cdb(tmp2.502,tmp2)
   eliminate_kronecker  (tmp2)  # cdb(tmp2.503,tmp2)
   tmp2 = flatten_Rabcd (tmp2)
   canonicalise         (tmp2)  # cdb(tmp2.506,tmp2)

   distribute           (tmp3)  # cdb(tmp3.501,tmp3)
   eliminate_metric     (tmp3)  # cdb(tmp3.502,tmp3)
   eliminate_kronecker  (tmp3)  # cdb(tmp3.503,tmp3)
   tmp3 = flatten_Rabcd (tmp3)
   canonicalise         (tmp3)  # cdb(tmp3.506,tmp3)

   distribute           (tmp4)  # cdb(tmp4.501,tmp4)
   eliminate_metric     (tmp4)  # cdb(tmp4.502,tmp4)
   eliminate_kronecker  (tmp4)  # cdb(tmp4.503,tmp4)
   tmp4 = flatten_Rabcd (tmp4)
   canonicalise         (tmp4)  # cdb(tmp4.506,tmp4)

   distribute           (tmp5)  # cdb(tmp5.501,tmp5)
   eliminate_metric     (tmp5)  # cdb(tmp5.502,tmp5)
   eliminate_kronecker  (tmp5)  # cdb(tmp5.503,tmp5)
   tmp5 = flatten_Rabcd (tmp5)
   canonicalise         (tmp5)  # cdb(tmp5.506,tmp5)

@@

   # raise the _{uv} indices to ^{ab}

   metric:= g^{a u} g^{b v} (  @(tmp0)
                              + (1/1) @(tmp1)
                              + (1/2) @(tmp2)
                              + (1/6) @(tmp3)
                              + (1/24) @(tmp4)
                              + (1/120) @(tmp5) ).  # cdb(metric.500,metric)

   distribute             (metric)   # cdb(metric.501,metric)
   eliminate_metric       (metric)   # cdb(metric.502,metric)
   eliminate_kronecker    (metric)   # cdb(metric.503,metric)
   metric = flatten_Rabcd (metric)   # cdb(metric.504,metric)
   canonicalise           (metric)   # cdb(metric.505,metric)

   substitute          (metric,$g_{a b} g^{b c} -> g_{a}^{c}$)
   substitute          (metric,$g_{b a} g^{b c} -> g_{a}^{c}$)
   substitute          (metric,$g_{b a} g^{c b} -> g_{a}^{c}$)
   substitute          (metric,$g_{a b} g^{c b} -> g_{a}^{c}$)
   eliminate_kronecker (metric)   # cdb(metric.506,metric)
   canonicalise        (metric)   # cdb(metric.507,metric)

   substitute (metric,$A^{a} -> x^{a}$)  # cdb (metric.508,metric)

   cdblib.create ('metric-inv.json')

   cdblib.put ('g^ab',metric,'metric-inv.json')

@@

   term0 = get_xterm (metric,0)   # cdb(term0.501,term0)
   term1 = get_xterm (metric,1)   # cdb(term1.501,term1)
   term2 = get_xterm (metric,2)   # cdb(term2.501,term2)
   term3 = get_xterm (metric,3)   # cdb(term3.501,term3)
   term4 = get_xterm (metric,4)   # cdb(term4.501,term4)
   term5 = get_xterm (metric,5)   # cdb(term5.501,term5)

   cdblib.put ('g^ab_0',term0,'metric-inv.json')
   cdblib.put ('g^ab_1',term1,'metric-inv.json')
   cdblib.put ('g^ab_2',term2,'metric-inv.json')
   cdblib.put ('g^ab_3',term3,'metric-inv.json')
   cdblib.put ('g^ab_4',term4,'metric-inv.json')
   cdblib.put ('g^ab_5',term5,'metric-inv.json')

@@

\begin{dgroup*}
   \begin{dmath*} \cdb*{term0.501} \end{dmath*}
   \begin{dmath*} \cdb*{term1.501} \end{dmath*}
   \begin{dmath*} \cdb*{term2.501} \end{dmath*}
   \begin{dmath*} \cdb*{term3.501} \end{dmath*}
   \begin{dmath*} \cdb*{term4.501} \end{dmath*}
   \begin{dmath*} \cdb*{term5.501} \end{dmath*}
\end{dgroup*}

@@

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp2.501} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.502} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.503} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.506} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp3.501} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.502} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.503} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.506} \end{dmath*}
\end{dgroup*}

@@

\begin{cadabra}
   Rterm0 := @(term0).
   Rterm1 := @(term1).  # zero
   Rterm2 := @(term2).
   Rterm3 := @(term3).
   Rterm4 := @(term4).
   Rterm5 := @(term5).

>>>>

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/metric.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{The metric tensor in Riemann normal coordinates}

In this notebook we compute the recursive sequences
\begin{align}
\label{eq:pdgab}
g_{ab,d\ue} &=  \left(g_{cb}\Gamma^{c}{}_{a(d}\right){}_{,\ue)}
               + \left(g_{ac}\Gamma^{c}{}_{b(d}\right){}_{,\ue)}\\[10pt]
\label{eq:pdGamma}
(n+3)\Gamma^a{}_{d(b,c\ue)} &= (n+1)\left(R^a{}_{(bc\Dot d,\ue)}
                               - \left(\Gamma^a{}_{f(c}\Gamma^f{}_{b{\Dot d}}\right){}_{,\ue)}\right)
\end{align}
for $n=1,2,3,\cdots$. Note in these equations that the (extended) index $\ue$ contains $n$
normal indices.

We then construct a Taylor series for the metric using
\begin{dmath*}[spread=5pt]
g_{ab}(x) = g_{ab} + g_{ab,c}x^c + \frac{1}{2!} g_{ab,cd}x^cx^d + \frac{1}{3!} g_{ab,cde}x^cx^dx^e + \cdots
          = g_{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g_{ab,\uc}\>x^{.\uc}
\end{dmath*}

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $g_{ab}$}

@@

\begin{dgroup*}
   \begin{dmath*} g_{ab,c} A^{c} = \cdb{term1.200} \end{dmath*}
   \begin{dmath*} g_{ab,cd} A^{c} A^{d} = \cdb{term2.200} \end{dmath*}
   \begin{dmath*} g_{ab,cde} A^{c} A^{d} A^{e} = \cdb{term3.200} \end{dmath*}
\end{dgroup*}

@@

\begin{dgroup*}
   \begin{dmath*} g_{ab,c} A^{c} = \cdb{term1.200} \end{dmath*}
   \begin{dmath*} g_{ab,cd} A^{c} A^{d} = \cdb{term2.303} \end{dmath*}
   \begin{dmath*} g_{ab,cde} A^{c} A^{d} A^{e} = \cdb{term3.305} \end{dmath*}
\end{dgroup*}

@@

\begin{dgroup*}
   \begin{dmath*} g_{ab,c} A^{c} = \cdb{term1.404} \end{dmath*}
   \begin{dmath*} g_{ab,cd} A^{c} A^{d} = \cdb{term2.404} \end{dmath*}
   \begin{dmath*} g_{ab,cde} A^{c} A^{d} A^{e} = \cdb{term3.403} \end{dmath*}
\end{dgroup*}

@@

\begin{align*}
   g_{ab}(x) &=   g_{ab}
                + g_{ab,c} x^c
                + \frac{1}{2!} g_{ab,cd} x^c x^d
                + \frac{1}{3!} g_{ab,cde} x^c x^d x^e +  \BigO{\eps^4}\\
             &= \cdb{metric4.501} + \BigO{\eps^4}
\end{align*}

@@

   def flatten_Rabcd (obj):
       substitute (obj,$R^{a}_{b c d}   -> g^{a e} R_{e b c d}$)
       substitute (obj,$R_{a}^{b}_{c d} -> g^{b e} R_{a e c d}$)
       substitute (obj,$R_{a b}^{c}_{b} -> g^{c e} R_{a b e d}$)
       substitute (obj,$R_{a b c}^{d}   -> g^{d e} R_{a b c e}$)
       unwrap     (obj)
       return obj

@@

@@

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $g_{ab}$}

@@

   # symmetrised partial derivatives of g_{ab}

   gab00:=g_{a b}.                                              # cdb (gab00.101,gab00)

   gab01:=g_{c b}\Gamma^{c}_{a d} + g_{a c}\Gamma^{c}_{b d}.    # cdb (gab01.101,gab01)

   gab02:=\partial_{e}{ @(gab01) }.                             # cdb (gab02.101,gab02)
   distribute   (gab02)                                         # cdb (gab02.102,gab02)
   product_rule (gab02)                                         # cdb (gab02.103,gab02)
   substitute   (gab02, $\partial_{d}{g_{a b}} -> @(gab01)$)    # cdb (gab02.104,gab02)
   distribute   (gab02)                                         # cdb (gab02.105,gab02)

   gab03:=\partial_{f}{ @(gab02) }.                             # cdb (gab03.101,gab03)
   distribute   (gab03)                                         # cdb (gab03.102,gab03)
   product_rule (gab03)                                         # cdb (gab03.103,gab03)
   substitute   (gab03, $\partial_{d}{g_{a b}} -> @(gab01)$)    # cdb (gab03.104,gab03)
   distribute   (gab03)                                         # cdb (gab03.105,gab03)

   gab04:=\partial_{g}{ @(gab03) }.                             # cdb (gab04.101,gab04)
   distribute   (gab04)                                         # cdb (gab04.102,gab04)
   product_rule (gab04)                                         # cdb (gab04.103,gab04)
   substitute   (gab04, $\partial_{d}{g_{a b}} -> @(gab01)$)    # cdb (gab04.104,gab04)
   distribute   (gab04)                                         # cdb (gab04.105,gab04)

   gab05:=\partial_{h}{ @(gab04) }.                             # cdb (gab05.101,gab05)
   distribute   (gab05)                                         # cdb (gab05.102,gab05)
   product_rule (gab05)                                         # cdb (gab05.103,gab05)
   substitute   (gab05, $\partial_{d}{g_{a b}} -> @(gab01)$)    # cdb (gab05.104,gab05)
   distribute   (gab05)                                         # cdb (gab05.105,gab05)

@@

\begin{cadabra}
   # prepare first six terms in the Taylor series expansion of g_{ab}(x)

@@

   end_stage_2 = time.time()

   # -------------------------------------------------------------------------------------------
   # this block of Xterms only produces formatted output, it's not part of the main computation
   # -------------------------------------------------------------------------------------------

@@

   metric:=@(term0)
         + (1/1) @(term1)  # zero
         + (1/2) @(term2)
         + (1/6) @(term3)
         + (1/24) @(term4)
         + (1/120) @(term5).  # cdb(metric.301,metric)

@@

   Xterm0 := @(term0).
   Xterm1 := (1/1) @(term1).
   Xterm2 := (1/2) @(term2).
   Xterm3 := (1/6) @(term3).
   Xterm4 := (1/24) @(term4).
   Xterm5 := (1/120) @(term5).

@@

   substitute (Xterm2,$g_{a b} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm2.301,Xterm2)
   substitute (Xterm3,$g_{a b} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm3.301,Xterm3)
   substitute (Xterm4,$g_{a b} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm4.301,Xterm4)
   substitute (Xterm5,$g_{a b} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm5.301,Xterm5)

   substitute (Xterm2,$g_{b a} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm2.301,Xterm2)
   substitute (Xterm3,$g_{b a} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm3.301,Xterm3)
   substitute (Xterm4,$g_{b a} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm4.301,Xterm4)
   substitute (Xterm5,$g_{b a} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm5.301,Xterm5)

   eliminate_metric (Xterm2)  # cdb(Xterm2.302,Xterm2)
   eliminate_metric (Xterm3)  # cdb(Xterm3.302,Xterm3)
   eliminate_metric (Xterm4)  # cdb(Xterm4.302,Xterm4)
   eliminate_metric (Xterm5)  # cdb(Xterm5.302,Xterm5)

   sort_product     (Xterm2)  # cdb(Xterm2.303,Xterm2)
   sort_product     (Xterm3)  # cdb(Xterm3.303,Xterm3)
   sort_product     (Xterm4)  # cdb(Xterm4.303,Xterm4)
   sort_product     (Xterm5)  # cdb(Xterm5.303,Xterm5)

   rename_dummies   (Xterm2)  # cdb(Xterm2.304,Xterm2)
   rename_dummies   (Xterm3)  # cdb(Xterm3.304,Xterm3)
   rename_dummies   (Xterm4)  # cdb(Xterm4.304,Xterm4)
   rename_dummies   (Xterm5)  # cdb(Xterm5.304,Xterm5)

   canonicalise     (Xterm2)  # cdb(Xterm2.305,Xterm2)
   canonicalise     (Xterm3)  # cdb(Xterm3.305,Xterm3)
   canonicalise     (Xterm4)  # cdb(Xterm4.305,Xterm4)
   canonicalise     (Xterm5)  # cdb(Xterm5.305,Xterm5)

   # push upper index to the left
   def tidy_Rabcd (obj):
       substitute (obj,$R_{a b c}^{d} -> - R^{d}_{c a b}$)
       substitute (obj,$R_{a b}^{c}_{d} -> R^{c}_{d a b}$)
       substitute (obj,$R_{a}^{b}_{c d} -> - R^{b}_{a c d}$)
       return obj

   Xterm0 = tidy_Rabcd (Xterm0)  # cdb(Xterm0.666,Xterm0)
   Xterm2 = tidy_Rabcd (Xterm2)  # cdb(Xterm2.666,Xterm2)
   Xterm3 = tidy_Rabcd (Xterm3)  # cdb(Xterm3.666,Xterm3)
   Xterm4 = tidy_Rabcd (Xterm4)  # cdb(Xterm4.666,Xterm4)
   Xterm5 = tidy_Rabcd (Xterm5)  # cdb(Xterm5.666,Xterm5)

   Xterm0 = reformat_xterm (Xterm0,  1)    # cdb(Xterm0.301,Xterm0)
   Xterm2 = reformat_xterm (Xterm2,  3)    # cdb(Xterm2.301,Xterm2)
   Xterm3 = reformat_xterm (Xterm3,  6)    # cdb(Xterm3.301,Xterm3)
   Xterm4 = reformat_xterm (Xterm4,360)    # cdb(Xterm4.301,Xterm4)
   Xterm5 = reformat_xterm (Xterm5,180)    # cdb(Xterm5.301,Xterm5)

   # canonicalise from reformat_xterm will slide upper index from left hand side
   # so now we slide the upper index back to the left

   Xterm0 = tidy_Rabcd (Xterm0)  # cdb(Xterm0.667,Xterm0)
   Xterm2 = tidy_Rabcd (Xterm2)  # cdb(Xterm2.667,Xterm2)
   Xterm3 = tidy_Rabcd (Xterm3)  # cdb(Xterm3.667,Xterm3)
   Xterm4 = tidy_Rabcd (Xterm4)  # cdb(Xterm4.667,Xterm4)
   Xterm5 = tidy_Rabcd (Xterm5)  # cdb(Xterm5.667,Xterm5)

   # metric to 3rd, 4th, 5th and 6th order terms in powers of x^a

@@

@@

\begin{dgroup*}
   \begin{dmath*} g_{ab}(x) = \cdb{Metric3.301} \end{dmath*}
   \begin{dmath*} g_{ab}(x) = \cdb{Metric4.301} \end{dmath*}
   \begin{dmath*} g_{ab}(x) = \cdb{Metric5.301} \end{dmath*}
   \begin{dmath*} g_{ab}(x) = \cdb{Metric6.301} \end{dmath*}
\end{dgroup*}

@@

\begin{cadabra}
   beg_stage_4 = time.time()
   # final housekeeping

   term1 = flatten_Rabcd (term1)         # cdb(term1.501,term1)
   term2 = flatten_Rabcd (term2)         # cdb(term2.501,term2)
   term3 = flatten_Rabcd (term3)         # cdb(term3.501,term3)
   term4 = flatten_Rabcd (term4)         # cdb(term4.501,term4)
   term5 = flatten_Rabcd (term5)         # cdb(term5.501,term5)

   eliminate_metric (term1)
   eliminate_metric (term2)
   eliminate_metric (term3)
   eliminate_metric (term4)
   eliminate_metric (term5)

   eliminate_kronecker (term1)
   eliminate_kronecker (term2)
   eliminate_kronecker (term3)
   eliminate_kronecker (term4)
   eliminate_kronecker (term5)

   sort_product (term1)
   sort_product (term2)
   sort_product (term3)
   sort_product (term4)
   sort_product (term5)

   rename_dummies (term1)
   rename_dummies (term2)
   rename_dummies (term3)
   rename_dummies (term4)
   rename_dummies (term5)

   canonicalise (term1)                  # cdb(term1.502,term1)
   canonicalise (term2)                  # cdb(term2.502,term2)
   canonicalise (term3)                  # cdb(term3.502,term3)
   canonicalise (term4)                  # cdb(term4.502,term4)
   canonicalise (term5)                  # cdb(term5.502,term5)

@@

   metric:=@(term0)
         + (1/1) @(term1)
         + (1/2) @(term2)
         + (1/6) @(term3)
         + (1/24) @(term4)
         + (1/120) @(term5).             # cdb(metric.501,metric)

   substitute (metric,$A^{a} -> x^{a}$)  # cdb (metric.502,metric)

   cdblib.create ('metric.json')

   cdblib.put ('g_ab',metric,'metric.json')

@@

   term0 = get_xterm (metric,0)          # cdb(term0.503,term0)
   term1 = get_xterm (metric,1)          # cdb(term1.503,term1)
   term2 = get_xterm (metric,2)          # cdb(term2.503,term2)
   term3 = get_xterm (metric,3)          # cdb(term3.503,term3)
   term4 = get_xterm (metric,4)          # cdb(term4.503,term4)
   term5 = get_xterm (metric,5)          # cdb(term5.503,term5)

   cdblib.put ('g_ab_0',term0,'metric.json')
   cdblib.put ('g_ab_1',term1,'metric.json')
   cdblib.put ('g_ab_2',term2,'metric.json')
   cdblib.put ('g_ab_3',term3,'metric.json')
   cdblib.put ('g_ab_4',term4,'metric.json')
   cdblib.put ('g_ab_5',term5,'metric.json')

@@

\begin{dgroup*}
   \begin{dmath*} \cdb*{term2.501} \end{dmath*}
   \begin{dmath*} \cdb*{term2.502} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term3.501} \end{dmath*}
   \begin{dmath*} \cdb*{term3.502} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term4.501} \end{dmath*}
   \begin{dmath*} \cdb*{term4.502} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term5.501} \end{dmath*}
   \begin{dmath*} \cdb*{term5.502} \end{dmath*}
\end{dgroup*}

@@

\begin{dgroup*}
   \begin{dmath*} \cdb*{metric.501} \end{dmath*}
   \begin{dmath*} \cdb*{metric.502} \end{dmath*}
\end{dgroup*}

@@

\begin{cadabra}
   Xterm0 := @(term0).
   Xterm1 := @(term1).  # zero
   Xterm2 := @(term2).
   Xterm3 := @(term3).
   Xterm4 := @(term4).
   Xterm5 := @(term5).
